---
title: "Simulating the DE"
format: html
---

# Prepare the Environment

```{r}
library(readxl)
library(MASS)
library(ggdag)
library(dagitty)
library(tidyverse)
```

# Wrangle the Data

Extract the Digital Economy Real Gross Output by Industry"

Create two data sets, one aggregated to the sub-attribute level and the other as granular The data contains subtotals which inflates the totals, those need to be removed We also want to understand how the market shifts over time, so I am input that variable into the simulation

```{r}
# read in the data
de22 <- read_excel("DE22.xlsx") %>% 
    filter(
    Measurement == "Digital Economy Real Gross Output by Industry",
    # remove 0s
    Value > 0) 
  

# remove subtotals, which I've annotated with "_TOTAL" in excel
#this allows for analysis of each industry at a granular level
gi_output <-  de22 %>% 
   filter(!str_detect(MicroAttribute, regex("_TOTAL", ignore_case = TRUE)))



# To make this easier to digest, I want look at the data at the third level.
# The BEA categorizes industries  down to 5 levels, and it's not consistent between industries
#  I want aggregate industries down to the 3 level which I called "sub-attributes" 

# create the summary for each year
# first I would like remove my labels so I can aggregate the totals together
# i'll use this data again
sub_industry_output<-  de22 %>% 
    mutate(
      Subcategory = str_replace(Subcategory, "_.*", ""),
      MicroAttribute = str_replace(MicroAttribute, "_.*", ""),
      Attribute = str_replace(Attribute, "_.*", ""),
      SubAttribute = str_replace(SubAttribute, "_.*", "")) %>%
    # remove these sub totals,  as all observations fall in one of the other
    filter(Category != "All Industries" & Category != "Private industries",
           # remove the granular category keeping only the observations totaled at the sub-attribute level
           SubAttribute == MicroAttribute)

# these categories have subtotals which for ease I'll remove manually
levels_remove <- c(
  "Arts, entertainment, recreation, accommodation, and food services",
  "Information",
  "Educational services, health care, and social assistance",
  "Finance, insurance, real estate, rental, and leasing",
  "Government",
  "Manufacturing",
  "Professional and business services")

sub_industry_output <- sub_industry_output %>%
  filter(!SubAttribute %in% levels_remove) %>%
  arrange(Year) %>%
  group_by(Category, Subcategory, Attribute) %>%
  mutate(
    Yearly_Change = replace_na(Value - lag(Value), 0),  # Replace NA with 0 in one step
    YOY = round(coalesce(Yearly_Change / lag(Value), 0), 2)) %>%
  ungroup() %>% 
  # renaming these columns so they make more sense
  dplyr::select(
    "Sector" = Subcategory,
    "Industry" = Attribute,
    "Specialization" = Category,
    Year,
    Value,
    Yearly_Change,
    YOY) 

sub_industry_output$Output_type <-  "Actual"


# Just the yearly Totals
gi_yearly_growth <-  de22 %>%
  filter(Category == "All Industries") %>%  
  dplyr::select(Category, Value, Year) %>%
  arrange(Year) %>%
  mutate(
  Yearly_Change = replace_na(Value - lag(Value), 0),  # Replace NA with 0 in one step    
  YOY = round(coalesce(Yearly_Change / lag(Value), 0), 2))
    # renaming these columns so they make more sense
 

```

# Collect Statistics for Simulation

```{r}
sub_industry_ss <- sub_industry_output %>%
  group_by(Sector,Industry, Specialization) %>%
  summarise(
    Mean_Value = round(mean(Value), 2),
    STD_Value = round(sd(Value), 2),
    Mean_YOY = round(mean(YOY[Year != 2017]), 2),
    STD_YOY = round(sd(YOY[Year != 2017]), 2))

sub_industry_ss
```

# I definetly want to turn this into an interactive graph!

```{r}
sub_industry_output$Year <- as.factor(sub_industry_output$Year)

sub_industry_output %>%
  filter(Year != 2017) %>% 
# Scatter plot of growth rates (YOY) over years
ggplot( aes(x = Year, y = YOY, color = Specialization)) +
  geom_line(aes(group = Specialization), alpha = 0.5) + 
  theme_minimal() +
  labs(
    title = "Growth Rates (YOY) by Specialization Over Years",
    x = "Year",
    y = "YOY Growth Rate",
    color = "Specialization") +
  theme(legend.position="none")

```

#What type of growth are we experiencing? It looks stocatic

```{r}


hist(sub_industry_output$YOY, breaks = 6, col = "lightblue", main = "Histogram of Data",
     xlab = "Data", ylab = "Frequency")

```

Stochastic Growth model, growth is unpredictable, compounded growth

```{r}
# Monte Carlo simulation
perform_monte_carlo <- function(data, n = 1000) {
  # create a list of results
  results <- list()
  
  #  create simulations for each type of specialized industry
  for (i in 1:nrow(data)) {
    sector <- data$Sector[i]
    industry <- data$Industry[i]
    specialization <- data$Specialization[i]
 
    
    # simulate growth,
    # using the 5 year average and standard deviation, generate 1000 random numbers that fall within the Gaussian distribution
    # with only five years data it's hard to see a pattern in the growth rates
    # i could use just the averages
    # we could do two type of models
    sim_growth <- rnorm(n, mean = data$Mean_YOY[i], sd = data$STD_YOY[i])
    
    # repeat the process for the overall industry output, using the simulated growth. applying the growth rate 
    sim_value <- rnorm(n, mean = data$Mean_Value[i] * (1 + sim_growth), sd = data$STD_Value[i])

    
    # Create a data frame for the current specialization
    sub_industry_sims <- data.frame(
      Sector = sector,
      Industry = industry,
      Specialization = specialization,
      Value = sim_value,
      YOY = round(sim_growth, 2)) # just for interprebability, do not want to double apply growth
    
    # Append to the results list
    results[[i]] <- sub_industry_sims
  }
  
  # Combine all results into a single data frame
  final_results <- bind_rows(results)
  
  return(final_results)
}

# Perform Monte Carlo simulation
simulation_results <- perform_monte_carlo(sub_industry_ss)

# View the results
head(simulation_results)
```

```{r}
de23 <-  simulation_results %>% 
  group_by(Sector,Industry, Specialization) %>%
  summarise(
    Bottom_Fifth_Value = quantile(Value, probs = 0.05),
    Mean_Value = mean(Value),
    Top_Fifth_Value = quantile(Value, probs = 0.95),
    Bottom_Fifth_Growth = quantile(YOY, probs = 0.05),
    Mean_Growth = mean(YOY),
    Top_Fifth_Growth = quantile(YOY, probs = 0.95))
```

```{r}
de23_long <- 
  pivot_longer(
    cols = starts_with("wk"),
    names_to = c("Value", "YOY"),
    names_prefix = "wk",
    values_to = "rank",
    values_drop_na = TRUE
  )
```

```{r}
sub_industry_output$Output_type <-  "Actual"
sub_industry_output <- 
```

```{r}
giy_dag <- dagify(
  Output ~ Category + Year,
  YOY ~ Output)

# Plot the DAG

ggdag(giy_dag)
```

# summary statisitcs for the aggregated industries entire data set

```{r}
gross_total <-  de22 %>% 
  filter(Measurement == "Digital Economy Real Gross Output by Industry",
         MicroAttribute == "All Industries_TOTAL_MA")

summary(gross_total$Value)
summary(gross_industry_output$Value)
summary(gross_industry_output_totals$Value)
```

## what do the summary stats look like by category?

```{r}
# library(ggforce)
# 
# 
# # plot each category
# filtered_data <- gross_industry_output_totals %>%
#   filter(Category != "All Industries",
#          Year == 2022)
# 
# 
# num_categories <- length(unique(filtered_data$Category))
# 
# 
# for (i in 1:num_categories) {
#   plot <- filtered_data %>%
#     ggplot(aes(y = Value, color = Category)) +
#     geom_boxplot() +
#     facet_wrap_paginate(~ Category, scales = "free", ncol = 1, nrow = 1, page = i) +
#     theme_bw() +
#     theme(legend.position = "none")
# 
#   print(plot)
# }

```

histogram of industry total for 2022

```{r}
gross_industry_output22 <- gross_industry_output %>% 
  filter(Year == 2022)

hist(gross_industry_output22$Bil_Dols, breaks = 5, col = "lightblue", main = "Histogram of Data",
     xlab = "Data", ylab = "Frequency")
summary(gross_industry_output22$Value)
```

```{r}
gross_industry_totals22 <- gross_industry_output_totals %>% 
  filter(Year == 2022)

hist(gross_industry_totals22$Bil_Dols, breaks = 5, col = "lightblue", main = "Histogram of Data",
     xlab = "Data", ylab = "Frequency")
summary(gross_industry_totals22$Value)
```

```{r}
data<- gross_industry_totals22$Value


fit <- fitdistr(data, "normal")

mean_fit <- fit$estimate["mean"]
sd_fit <- fit$estimate["sd"]

fit
```

```{r}
x <- seq(min(data), max(data), length = 100)
normal_curve <- dnorm(x, mean = mean_fit, sd = sd_fit)
normal_density <- dnorm(x, mean = mean_fit, sd = sd_fit)
scaled_density <- normal_density * length(data) * diff(hist(data, breaks = 5)$breaks[1:2])
lines(x, scaled_density, col = "red", lwd = 2)
```

# export data for excel exercise

```{r}
write.csv(gross_industry_totals22, "gross_industry_totals22.csv")
```

# next generate future samples
